Discrete Quantum Theories and Computing

### 2019-01-16 dissertation.tex
dissertation.lyx "5.3.1 Finite-Precision Extension of Gleason's Theorem" and "Chapter 6 Further Discussion" Updating...

To that end, we introduce the concept of \emph{interval maps}, which
we can use to construct a consistent family of QIVPMs. An interval
map $f:\left[0,1\right]\rightarrow\mathscr{I}$ maps every real-valued
probability $x\in\left[0,1\right]$ to a set of intervals $f\left(x\right)=\left[\ell,r\right]$
containing $x$, where $\left[0,1\right]$ denotes the set of real-valued
probabilities (this should not be confused with the interval-valued
probability $\unknown$). We also need a notion of \emph{norm} to
quantify the uncertainty in measurements and the distance between
(pure or mixed) states. The norm of a collection of intervals $\mathscr{I}$,
$\left\Vert \mathscr{I}\right\Vert $, is defined by the maximum length
of intervals in it. The norm of a pure state $\rho=\proj{\psi}$ is
defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
For any given Hermitian operator~$A$, we choose the operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
this norm is the maximum absolute value of its eigenvalues. Then,
a finite-precision extension of Gleason's theorem can be stated as
follows.

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map~$f$
with a conventional quantum probability measure~$\muB_{\rho}$, and
the states consistent with the composition $f\circ\muB_{\rho}$ collapses
to a single state as the maximum length of intervals in $\mathscr{I}$,
$\left\Vert \mathscr{I}\right\Vert $, shrinks to $0$. In contrast,
the globally inconsistent QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
has the minimal sharp range~$\mathscr{I}_{0}$ with $\left\Vert \mathscr{I}_{0}\right\Vert =1$.
This suggests that even for general QIVPMs, shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize them. In another word, if $\left\Vert \mathscr{I}\right\Vert $
is shorter than a number, all QIVPMs map to $\mathscr{I}$ might have
an non-empty core.

@@

so does their interval-valued probabilities, i.e., $\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\necess-\bar{\mu}\left(\proj{\psi}\right)$.

@@

for $\ket{\psi}\in\CP{2}$. According to Lemma~\ref{thm:convex-3-1},

@@

if Eqs.~(\ref{eq:QIVPM-constraints}) and

@@

holds for any orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
By applying Eq.~(\ref{eq:QIVPM-complement}) on the left-hand side,

@@

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea how hard this question could
be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum probability
space is glued by a family of classical probability spaces. This is
similar to the situation that a manifold is glued by many local coordinates.
When each small piece has \emph{exactly} the same and positive curvature,
the Killing-Hopf theorem asserts this manifold is a sphere, but little
can we say even if the curvature has only small deviation from constant.
Similar situation might happen when approximating Gleason's theorem.
When the sum of $\frameL$ on each ``local'' classical probability
space, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$, is \emph{exactly}
the same and equal to $1$, Gleason's theorem can assert $\frameL$
can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}). However, when
each local classical probability space becomes imprecise, a general
$\frameL$ might be as wild as we can imagine, and we might need to
know a little bit more, like the function is of the form $f\circ\muB_{\rho}$,
to deduce its global property.

@@

After we build the quantum interval-valued probability models, we
might think how powerful a quantum computer could be based on this
model. Since the conventional quantum circuit model manipulates the
probability amplitudes of quantum states, to build a quantum computing
model above QIVPMs, we might either consider evolving the states in
the core of a QIVPM, or find a way to evolve QIVPMs directly. Both
strategies are not straightforward. As we proved in Thm.~\ref{thm:Non-extensible-of-Gleason's},
a QIVPM might have an empty core which cannot be evolved over time.
If we want to evolve and compute QIVPMs directly, we need to get the
QIVPM of the whole system from the QIVPM for each subsystem or qubit.
However, there is no straightforward way to define the tensor product
among QIVPMs because we cannot summarize a probability measure into
a state as we did in conventional quantum theory.

>>>>

To that end, we introduce the concept of \emph{interval maps}, which
we can use to construct a consistent family of QIVPMs. An interval
map $f:\left[0,1\right]\rightarrow\mathscr{I}$ maps every real-valued
probability $x\in\left[0,1\right]$ to a set of intervals $f\left(x\right)=\left[\ell,r\right]$
containing $x$, where $\left[0,1\right]$ denotes the set of real-valued
probabilities (this should not be confused with the interval-valued
probability $\unknown$). We also need a notion of \emph{norm} to
quantify the uncertainty in measurements and the distance between
(pure or mixed) states. The norm of a collection of intervals $\mathscr{I}$,
$\left\Vert \mathscr{I}\right\Vert $, is defined as the maximum length
of intervals in it. The norm of a pure state $\rho=\proj{\psi}$ is
defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
For any given Hermitian operator~$A$, we choose the operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
this norm is the maximum absolute value of its eigenvalues. Then,
a finite-precision extension of Gleason's theorem can be stated as
follows.

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map
with a conventional quantum probability measure, and the states consistent
with the composite QIVPM collapses to a single state as the maximum
length of intervals in $\mathscr{I}$, $\left\Vert \mathscr{I}\right\Vert $,
shrinks to $0$. In contrast, the globally inconsistent QIVPM defined
in Table~\ref{tab:non-Born-QIVPM} has the least sharp range~$\mathscr{I}_{0}$
with $\left\Vert \mathscr{I}_{0}\right\Vert =1$. This suggests a
possibility that shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize general QIVPMs, and when $\left\Vert \mathscr{I}\right\Vert $
is shorter than a number, all QIVPMs map to $\mathscr{I}$ might have
non-empty cores.

@@

so does their interval-valued probabilities, i.e., 
\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\necess-\bar{\mu}\left(\proj{\psi}\right)\,.\label{eq:QIVPM-complement-states}

@@

for any irreducible state $\ket{\psi}\in\CP{2}$. According to Lemma~\ref{thm:convex-3-1},

@@

if $\bar{\mu}$ satisfies Eqs.~(\ref{eq:QIVPM-constraints}) and

@@

for all orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
By applying $\proj{\psi_{1}}+\proj{\psi_{2}}=\mathbb{1}-\proj{\psi_{0}}$
and Eq.~(\ref{eq:QIVPM-complement-states}) on the left-hand side,

@@

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea how hard this question could
be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum probability
space is glued by a family of classical probability spaces. This is
similar to the situation that a manifold is glued by many local coordinates.
When each small piece has \emph{exactly} the same and positive curvature,
the Killing-Hopf theorem asserts this manifold is a sphere, but little
can we say even if the curvature has only small deviation from constant.
Similar situation might happen when approximating Gleason's theorem.
When the sum of $\frameL$ on each ``local'' classical probability
space, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$, is \emph{exactly}
the same and equal to $1$, Gleason's theorem assert that $\frameL$
can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}). However, when
each local classical probability space becomes imprecise, a general
$\frameL$ might be as wild as we can imagine, and we might need to
know a little bit more, like its QIVPM is a composite function, to
deduce its global property.

@@

After we build the quantum interval-valued probability models, we
might think how powerful a quantum computer could be based on this
model. Since the conventional quantum circuit model manipulates the
probability amplitudes of quantum states, to build a quantum computing
model above QIVPMs, we might consider either manipulating the states
in the core of a QIVPM, or finding a way to manipulate QIVPMs directly.
However, both strategies are not straightforward. As we proved in
Thm.~\ref{thm:Non-extensible-of-Gleason's}, a QIVPM might have an
empty core which cannot be evolved over time. If we want to manipulate
and compute QIVPMs directly for multi-qubit algorithms, we need to
glue the QIVPM for each qubit or subsystem together to get the QIVPM
of the whole system, and this is not straightforward either.

A successful interval-valued theory might be further extended to the
quantum theory over finite fields, and consider the following quantum
interval-valued probability measures over finite fields.

\begin{definition}[Quantum Interval-valued Probability Measures
over Finite Fields]\label{def:QIVPMFF}Given a vector space~$\mathcal{H}$
of dimension~$D$ over the complexified field~$\Fpp$, the set of
events~$\events_{p^{2}}$ is defined as in Def.~\ref{def:QPMFF},
and $\mathscr{I}$ is a collection of intervals. Then, a quantum interval-valued
probability measure over finite field $\bar{\mu}:\events_{p^{2}}\rightarrow\mathscr{I}$
assigns a probability to each event (projection operator~$P$) subject
to $\bar{\mu}(\mathbb{0})=\imposs$, $\bar{\mu}(\mathbb{1})=\necess$,
$\bar{\mu}\left(\mathbb{1}-P\right)=\necess-\bar{\mu}\left(P\right)$,
and satisfying for each pair of \emph{commuting} projectors~$P_{0}$
and~$P_{1}$ with $P_{0}P_{1}=P_{1}P_{0}$, 
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.\label{eq:QuantumInterval-valuedProbability-Inclusion-Finite-Field}
\end{equation}
\end{definition}

\noindent Understanding the properties of these probability measures
and whether we could define any ``sensible'' Born upon them will
be the next natural extension for seeking a computable quantum theory.
